---
title: '`r paste0("Latest Statewide COVID-19 Reproductive Number Estimates and Projections for ", TEMPLATE_STATE_NAME)`'
author: "Grant Brown, Josh Tomiyama"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, echo = FALSE, results='hide'}
suppressWarnings(
  suppressPackageStartupMessages({
    library(knitr)
    library(dplyr)
    library(ggplot2)
    library(kableExtra )
    library(gridExtra)
    library(ABSEIR)
    library(plotly)
    library(htmltools)
    })
  )

resultDirectory <- "../../Results/"
if (!dir.exists(resultDirectory)){
  stop("\nYou are running this file in the wrong directory\n")
}
resultDir <- setdiff(dir(resultDirectory), "Readme")
## Determine if we're supposed to be working in a specific folder, or grabbing the latest

if (!exists("resultsDirectoryName")){
  maxidx <- which.max(as.Date(resultDir))
  rprtDate <- max(as.Date(resultDir), na.rm = TRUE)
  resultsDirectoryName <- paste0(resultDirectory, resultDir[maxidx])
} else{
  rprtDate <- as.Date(resultsDirectoryName)
  resultsDirectoryName <- paste0(resultDirectory, resultsDirectoryName)
}
maxPredDate <- max( rprtDate + 30, as.Date("2020-06-01"))
clean_idx <- ifelse(TEMPLATE_STATE_IDX < 10, 
                    paste0("0", trimws(TEMPLATE_STATE_IDX)), 
                    TEMPLATE_STATE_IDX)
rsltFileIndicator <- grepl(clean_idx, substr(dir(resultsDirectoryName), 1, 2)) &
                     grepl(paste0("[",paste0(
                       TEMPLATE_INTERVENTION_TYPES,collapse = ""),"]"),
                           substr(dir(resultsDirectoryName), 4, 4))

if(sum(rsltFileIndicator) > 0){
  resultFiles <- paste0(resultsDirectoryName,"/", 
                        dir(resultsDirectoryName)[rsltFileIndicator])
}else{
  stop("\nResults files were not found for", TEMPLATE_STATE_IDX, "\n")
}

analysisResults <- lapply(resultFiles, function(rf){
  e <- new.env()
  load(rf, envir = e)
  e
})
```

```{r,echo=FALSE, results='asis'}
cat(paste0("<h1>Using report compiled on: ", rprtDate, "</h1>\n"))
set.seed(as.numeric(rprtDate))
outputPath <- paste0("../../StateOutput/", rprtDate, "/")
if(!dir.exists(outputPath)){
  dir.create(outputPath)
}
```

## Introduction

Scientific questions of interest:

* Do stay at home orders affect the number of cases of COVID-19?

### Reproductive Numbers

Interventions to control infectious diseases strive to drive the "reproductive number" below 1. While this quantity comes in many forms, and is derived in many ways for different models, in general it captures the number of secondary infections expected per infectious individual. When this number falls below 1, the spread of the disease is expected to slow and stop, while values greater than 1 indicate that the disease is likely to continue to spread exponentially until some other phenomenon interrupts transmission (additional interventions, natural behavioral changes, a lack of susceptible individuals). 

The plot below provides the latest plot estimating our relationship to this threshold. Namely, it estimates the probability that the reproductive number has fallen below 1 at any point during the outbreak, and provides a forecast of the number about 1 month into the future.

## Methods

Typically, we do not know the functional form of how a public policy will affect the probability of being exposed to an infectious disease. Meaning, we do not know whether the effect is linear, quadratic, or exponential with time. To account for this functional uncertainty, we consider a set of 3 models and take a weighted average over them to conduct statistical inference. Assuming that the estimates from these models are not overtly bias, this model averaging strategy should help reduce the immense variability inherent in acute infectious disease outbreak models.

Model 1 assumes an instantaneous shift in (a function of) the probability of exposure once the public policy order starts and again when it stops. Model 2 assumes a piecewise linear effect with time. Model 3 specifies a basis spline expansion over the mandate time. For states without a stay at home order, a basis spline expansion on the observed epidemic time are used. Models 4 and 5 are based on google mobility data and are not directly linked to the stay at home order intervention dates, but one would changes in mobility in response to the stay at home orders. Model 4 is based on workplace mobility whereas model 5 is based on retail and recreation mobility. See the results section for visualizations of these model intervention effects.

## Data

Models are fit using the publicly available New York Times Covid-19 data. 

## Model Summary and Diagnostics {.tabset}

### Model Comparisons

```{r}
intervention_types <- substr(dir(resultsDirectoryName)[rsltFileIndicator], 
                             4, 
                             4)
```

Intervention Model Types Used: `r intervention_types`

```{r results='asis'}
term_eps <- NULL
mods <- list()
for(i in 1:length(intervention_types)){
  term_eps <- c(term_eps, analysisResults[[i]]$single_state_results$result$modelObject$current_eps)
  mods[[i]] <- analysisResults[[i]]$single_state_results$result$modelObject
}

### Table reads row is x times better than column
eps_df <- data.frame(inter_type = paste0("Intervention: ",
                                 intervention_types),
             term_eps = term_eps)

if(length(mods) > 1){
  cat("\n\n The Bayes Factor is a measurement of how much evidence is in favor of one model over the other. In the table below, one can read the table as the model in row i has 'x' times more evidence in favor of it over the model in column j.\n\n")
  
  invisible(bf <- compareModels(mods, 
                                n_samples = 1000, 
                                batch_size = 2000))
  colnames(bf) <- rownames(bf) <- paste0("Intervention", intervention_types)
  bf %>% kable(digits = 3,
             align = c("c"), 
             caption = "Model Bayes Factors") %>% 
  kable_styling(bootstrap_option = c("striped", "condensed"), 
                full_width = FALSE) 
}else{
  bf <- NA
}

cat("\n\nBelow are the terminating epsilons for each intervention model. For reference, if a model whose mortality predictions were off by a count of 10 at every time point the terminating epsilon would be ~",
    
    round(sqrt(10^2*nrow(analysisResults[[1]]$single_state_results$data))),
    "and for a model whose predictions were off by a count of 20 it would be ~",
    round(sqrt(20^2*nrow(analysisResults[[1]]$single_state_results$data)))
    )

eps_df %>% kable(digits = 3,
             align = c("l","c"), 
             caption = "Model Epsilons", 
             col.names = c("Intervention Type", "Terminating Epsilon")) %>% 
  kable_styling(bootstrap_option = c("striped", "condensed"), 
                full_width = FALSE)

if(length(intervention_types) == 1 & (0 %in% intervention_types)){
  InterventionStartDate <- analysisResults[[1]]$opt$date
  InterventionEndDate <- analysisResults[[1]]$opt$reopen
}else{
  InterventionStartDate <- InterventionEndDate <- NA
}

eps_df <- eps_df %>% mutate(state_idx = TEMPLATE_STATE_IDX, 
                            state = TEMPLATE_STATE_NAME,
                            start_date = InterventionStartDate,
                            end_date = InterventionEndDate)
save(eps_df, 
     bf,
     file = paste0(outputPath, TEMPLATE_STATE_IDX, "_model_eps.rda")
     )

```

```{r}

### x should be a SpatialSEIR summary object
htmlSummary <- function(x,...){
    cat(
    "
    <strong>Summary: SEIR Model</strong>\n\n
    <table>
      <tr>
      <td><strong> Number Locations </strong></td>
      <td> &nbsp; </td>
      <td>",x$nLoc,"</td>
      </tr>
      <tr>
      <td><strong> Number Time Points </strong></td>
      <td> &nbsp; </td>
      <td>", x$nTpt,"</td>
      </tr>
      <tr>
      <td><strong> Data Model Parameters: </strong></td>
      <td> &nbsp; </td>
      <td>",x$dataModelParams,"</td>
      </tr>
      <tr>
      <td><strong> Exposure Process Parameters: </strong></td>
      <td> &nbsp; </td>
      <td>",x$exposureParams ,"</td>
      </tr>
      <tr>
      <td><strong> Reinfection Model Parameters: </strong></td>
      <td> &nbsp; </td>
      <td>", x$reinfectionParams, "</td>
      </tr>
      <tr>
      <td><strong> Spatial Parameters: </strong></td>
      <td> &nbsp; </td>
      <td>", x$spatialParams ,"</td>
      </tr>
      <tr>
      <td><strong> Transition Parameters: </strong></td>
      <td> &nbsp; </td>
      <td>", x$transitionParams,"</td>
      </tr>
    </table>
    "
    )
    cat("\n\n")
    kable(x$parameterEstimates, 
          digits = 3, 
          align = c("c"), 
          caption = "Parameter Estimates",
          format.args = list(big.mark = ",")) %>% 
    kable_styling(bootstrap_option = c("striped", "condensed"), 
                full_width = FALSE) %>% print()
}
```

```{r}
##### Create all necessary Plots
IplotList <- list()
RplotList <- list()

### obs_data here should depend on whether there is data past the analysis date...
dtobj <- analysisResults[[1]]$single_state_results$data$date
obs_data <- analysisResults[[1]]$single_state_results$data %>%
  select(date) %>%
  left_join(analysisResults[[1]]$stateDataObserved) %>%
  filter(date <= max(dtobj))
obs_I <- obs_data$casesNonCum
obs_R <- obs_data$deathsNonCum

for(i in 1:length(intervention_types)){
  sims <- analysisResults[[i]]$single_state_results$result$simulationResults
  
  ## Posterior Predictive Distribution Plots
  cases <- Reduce(cbind, lapply(sims, function(x){x$I_star}))
  
  ### I_star Plot
  pq <- c(0.1, 0.25, 0.5,0.75,0.9)
  qcases <- apply(cases, 1, quantile, probs = pq)
  rownames(qcases) <- paste0("q", pq*100)
  out <- cbind(data.frame(date = dtobj, obs_cases=obs_I),
               t(as.data.frame(qcases)))
  outFilt <- out[out$date <= maxPredDate,]
  Iplot <- ggplot(data = outFilt, aes(x=date, y = q50)) + 
  geom_line(color="blue") + 
  geom_vline(xintercept = as.Date(rprtDate)) + 
  geom_ribbon(aes(ymin = q10, ymax=q90), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = q25, ymax=q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  xlab("Date") + 
  ylab("Daily Case Count") +
  ggtitle("Predicted Case Counts") + 
  ggtitle("Predicted Case Counts - Current Estimates and Projections\n 50% and 80% Intervals") +  
  geom_point(aes(x = date, y = obs_cases))
  IplotList[[i]] <- ggplotly(Iplot)
  
  ### R_star Plot
  pred_deathsNonCum <- Reduce(cbind, lapply(1:length(sims), function(j){
    pi_R <- analysisResults[[i]]$single_state_results$result$params[j,]
    pi_R <- pi_R[names(pi_R) == "report_fraction"]
    deaths <- sims[[j]]$R_star*pi_R
    deaths
  }))
  pq <- c(0.1, 0.25, 0.5,0.75,0.9)
  qdeaths <- apply(pred_deathsNonCum, 1, quantile, probs = pq)
  rownames(qdeaths) <- paste0("q", pq*100)
  out <- cbind(data.frame(date = dtobj, obs_cases = obs_R),
               t(as.data.frame(qdeaths)))
  outFilt <- out[out$date <= maxPredDate,]
  Rplot <- ggplot(data = outFilt, aes(x=date, y = q50)) + 
  geom_line(color="blue") + 
  geom_vline(xintercept = as.Date(rprtDate)) + 
  geom_ribbon(aes(ymin = q10, ymax=q90), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = q25, ymax=q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  xlab("Date") + 
  ylab("Daily Death Count") +
  ggtitle("Projected Mortality Counts") + 
  ggtitle("Projected Mortality Counts - Current Estimates and Projections\n 50% and 80% Intervals") +  
  geom_point(aes(x = date, y = obs_cases))
  RplotList[[i]] <- ggplotly(Rplot)
}
```

```{r results='asis'}
#### Generate Markdown code
for(i in 1:length(intervention_types)){
  cat("### Intervention Model Type", intervention_types[i],"\n\n")
  
  cat("#### Summary of Parameters:\n\n")
  htmlSummary(summary(mods[[i]]))
  cat("\n\n")

  cat("#### Posterior Predictive Distributions:\n\n")
  
  cat("\nBelow is the predicted Deaths per day based on this model. We would expect and hope the model to perform relatively well as the models are trained using mortality data. The blue line represents the median of our predictive distribution while the points represent the observed case counts.\n\n")
  cat(renderTags(RplotList[[i]])$html)
  cat("\n\n")
  
  cat("Below is the predicted new cases per day based on this model. We would expect the some underperformance in the model as the models are trained using mortality data. The blue line represents the median of our predictive distribution while the points represent the observed case counts.\n")
  cat("\n\n")
  cat(renderTags(IplotList[[i]])$html)
  cat("\n\n")
  
}
```

```{r echo=FALSE, messages=FALSE, warning=FALSE}
# attach the Dependencies
# since they do not get included with renderTags(...)$html
deps <- lapply(
  RplotList,
  function(x){
    renderTags(x)$dependencies
  }
)

attachDependencies(
  tagList(),
  unlist(deps,recursive=FALSE)
)

deps2 <- lapply(
  IplotList,
  function(x){
    renderTags(x)$dependencies
  }
)

attachDependencies(
  tagList(),
  unlist(deps,recursive=FALSE)
)
```

## Model-Averaged Results {.tabset}


### Latest EA-RN information for `r cat(TEMPLATE_STATE_NAME)` 


```{r, echo=FALSE}
R0frame <- Reduce("cbind", lapply(1:length(analysisResults), function(i){
  sims <- analysisResults[[i]]$single_state_results$result$simulationResults
  
  reaOut <- Reduce("cbind", lapply(1:length(sims), function(j){
    REA <- sims[[j]]$R_EA
    REA
  }))
  
}))
## Each eps is the epsilon associated with the data produced from that specific particle (wording)
eps <- Reduce("c", lapply(1:length(analysisResults), function(i){
  analysisResults[[i]]$single_state_results$result$modelObject$epsilon
}))
invEps <- 1/eps
invEps <- invEps/sum(invEps)
sampIdx <- sample(1:length(invEps), size = 1000, prob = invEps, replace = TRUE)
R0samp <- R0frame[,sampIdx]
dtobj <- analysisResults[[1]]$single_state_results$data$date
R0bounds <- apply(R0samp, 1, quantile, probs = c(0.1, 0.25, 0.5,0.75,0.9))
out <- data.frame(date = dtobj, q1 = R0bounds[1,], 
                  q25 = R0bounds[2,],
                  q5 = R0bounds[3,], 
                  q75 = R0bounds[4,],
                  q9 = R0bounds[5,],
                  Observed =  analysisResults[[1]]$single_state_results$data$deaths)
outFilt <- out[out$date <= maxPredDate,]
p1 <- ggplot(data = outFilt, aes(x=date, y = q5)) + 
  geom_line(color="blue") + 
  geom_vline(xintercept = as.Date("2020-06-01")) + 
  geom_ribbon(aes(ymin = q1, ymax=q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = q25, ymax=q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("R-EA") + xlab("Date") + 
  ggtitle("Reproductive Number")+ 
  ggtitle("Reproductive Number - Current Estimates and Projections\n 50% and 80% Intervals")+  
  geom_hline(yintercept = 1,linetype = "dashed", color = "black", size = 1)+
  theme_light()
p1
outFilt <- outFilt %>% mutate(state_idx = TEMPLATE_STATE_IDX, state = TEMPLATE_STATE_NAME)
save(outFilt, file = paste0(outputPath, TEMPLATE_STATE_IDX, "_R0.rda"))
```

### Latest Projected Mortality

```{r, echo = FALSE}
## Get mortality projections 
Mortalityframe <- Reduce("cbind", lapply(1:length(analysisResults), function(i){
  sims <- analysisResults[[i]]$single_state_results$result$simulationResults
  
YOut <- Reduce("cbind", lapply(1:length(sims), function(j){
    pi_R <- analysisResults[[i]]$single_state_results$result$params[j,]
    pi_R <- pi_R[names(pi_R) == "report_fraction"]
    deaths <- sims[[j]]$R_star*pi_R
    deaths
  }))
  YOut
}))
eps <- Reduce("c", lapply(1:length(analysisResults), function(i){
  analysisResults[[i]]$single_state_results$result$modelObject$epsilon
}))
invEps <- 1/eps
invEps <- invEps/sum(invEps)
sampIdx <- sample(1:length(invEps), size = 1000, prob = invEps, replace = TRUE)
MortalityframeSamp <- Mortalityframe[,sampIdx]
MortalityframeSamp <- apply(MortalityframeSamp, 2, cumsum)
dtobj <- analysisResults[[1]]$single_state_results$data$date
may1Idx <- which(dtobj == "2020-05-01")
mortalityQuantiles <- apply(MortalityframeSamp, 1, quantile, 
                            probs = c(0.1,0.25, 0.5,0.75,0.9))
out <- data.frame(date = dtobj, q1 = mortalityQuantiles[1,], 
                  q25 = mortalityQuantiles[2,],
                  q5 = mortalityQuantiles[3,], 
                  q75 = mortalityQuantiles[4,],
                  q9 = mortalityQuantiles[5,],
                  Observed =  analysisResults[[1]]$single_state_results$data$deaths)
outFilt <- out[out$date <= maxPredDate,]
outFilt.latest <- outFilt
p1 <- ggplot(data = outFilt, aes(x=date, y = q5)) + 
  geom_vline(xintercept = as.Date("2020-06-01")) +
  geom_point(x = outFilt$date, y = outFilt$Observed) + 
  geom_line(color="blue") + 
  geom_ribbon(aes(ymin = q1, ymax=q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = q25, ymax=q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("Cumulative Mortality") + xlab("Date") + 
  ggtitle("Six Week Cumulative Mortality")+ 
 
  ggtitle("Projected Cumulative Mortality\n 50% and 80% Intervals") +  
  theme_light()
suppressWarnings(
  print(p1)
)
p2 <- ggplot(data = outFilt, aes(x=date, y = q5))+  scale_y_continuous(trans='log10') + 
  geom_line(color="blue") + 
  geom_vline(xintercept = as.Date("2020-06-01")) +
  geom_point(x = outFilt$date, y = log10(outFilt$Observed) )+
  geom_ribbon(aes(ymin = outFilt$q1, ymax=outFilt$q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = outFilt$q25, ymax=outFilt$q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("Cumulative Mortality") + xlab("Date") + 
  ggtitle("Projected Cumulative Mortality\n (log10 scale)")+  
  theme_light() 
suppressWarnings(
  print(p2)
)
out.joined <- out  %>% mutate(CIprnt = paste0(round(q5), " (", round(q1), ", ", round(q9), ")"))
out.ToPrint <- select(out.joined, date, Observed, CIprnt) %>% filter(date <= maxPredDate)
colnames(out.ToPrint) <- c("Date", "Observed Mortality", "Predicted Mortality")
kable(out.ToPrint, digits = 1,align = c("c"), caption = "Mortality: Latest") %>% 
  kable_styling(bootstrap_option = c("striped", "condensed"), 
                full_width = FALSE)
outFilt <- outFilt %>% mutate(state_idx = TEMPLATE_STATE_IDX, state = TEMPLATE_STATE_NAME)
save(outFilt, file = paste0(outputPath, TEMPLATE_STATE_IDX, "_CumMort.rda"))
 
```

### Latest Cumulative Case Count

```{r, echo = FALSE}
## Get Incidence projections 
Incidenceframe <- Reduce("cbind", lapply(1:length(analysisResults), function(i){
  sims <- analysisResults[[i]]$single_state_results$result$simulationResults
  YOut <- Reduce("cbind", lapply(1:length(sims), function(j){
      cases <- sims[[j]]$I_star
      cases
    }))
  YOut
}))


eps <- Reduce("c", lapply(1:length(analysisResults), function(i){
  analysisResults[[i]]$single_state_results$result$modelObject$epsilon
}))
invEps <- 1/eps
invEps <- invEps/sum(invEps)
sampIdx <- sample(1:length(invEps), size = 1000, prob = invEps, replace = TRUE)
IncidenceframeSamp <- Incidenceframe[,sampIdx]
IncidenceframeSamp <- apply(IncidenceframeSamp, 2, cumsum)
dtobj <- analysisResults[[1]]$single_state_results$data$date
may1Idx <- which(dtobj == "2020-05-01")
IncidenceQuantiles <- apply(IncidenceframeSamp, 1, quantile, probs = c(0.1,0.25, 0.5,0.75,0.9))
out <- data.frame(date = dtobj, 
                  q1 = IncidenceQuantiles[1,], 
                  q25 = IncidenceQuantiles[2,],
                  q5 = IncidenceQuantiles[3,], 
                  q75 = IncidenceQuantiles[4,],
                  q9 = IncidenceQuantiles[5,],
                  Observed =  analysisResults[[1]]$single_state_results$data$cases)
outFilt <- out[out$date <= maxPredDate,]
outFilt.latest <- outFilt
p1 <- ggplot(data = outFilt, aes(x=date, y = q5)) + 
  geom_vline(xintercept = as.Date("2020-06-01")) +
  geom_point(x = outFilt$date, y = outFilt$Observed) + 
  geom_line(color="blue") + 
  geom_ribbon(aes(ymin = q1, ymax=q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = q25, ymax=q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("Cumulative Incidence") + xlab("Date") + 
  ggtitle("Six Week Cumulative Cases")+ 
 
  ggtitle("Projected Cumulative Cases\n 50% and 80% Intervals") +  
  theme_light()
suppressWarnings(
  print(p1)
)
p2 <- ggplot(data = outFilt, aes(x=date, y = q5))+  scale_y_continuous(trans='log10') + 
  geom_line(color="blue") + 
  geom_vline(xintercept = as.Date("2020-06-01")) +
  geom_point(x = outFilt$date, y = log10(outFilt$Observed) )+
  geom_ribbon(aes(ymin = outFilt$q1, ymax=outFilt$q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = outFilt$q25, ymax=outFilt$q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("Cumulative Incidence") + xlab("Date") + 
  ggtitle("Projected Cumulative Incidence\n (log10 scale)")+  
  theme_light() 
suppressWarnings(
  print(p2)
)
out.joined <- out  %>% mutate(CIprnt = paste0(round(q5), " (", round(q1), ", ", round(q9), ")"))
out.ToPrint <- select(out.joined, date, Observed, CIprnt) %>% filter(date <= maxPredDate)
colnames(out.ToPrint) <- c("Date", "Observed Incidence", "Predicted Incidence")
kable(out.ToPrint, digits = 1,align = c("c"), caption = "Incidence: Latest") %>% 
  kable_styling(bootstrap_option = c("striped", "condensed"), 
                full_width = FALSE) 
outFilt <- outFilt %>% mutate(state_idx = TEMPLATE_STATE_IDX, 
                              state = TEMPLATE_STATE_NAME)
save(outFilt, file = paste0("../../StateOutput/",
                            TEMPLATE_STATE_IDX, "_CumCases.rda"))
 
```



