---
title: '`r paste0("Latest Statewide COVID-19 Reproductive Number Estimates and Projections for ", TEMPLATE_STATE_NAME)`'
author: "Grant Brown"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo = FALSE}
suppressMessages({
library(knitr)
library(dplyr)
library(ggplot2)
library(kableExtra )
library(gridExtra)
})
resultDirectory <- "../Results/"
if (!dir.exists(resultDirectory)){
  stop("\nYou are running this file in the wrong directory\n")
}
resultDir <- setdiff(dir(resultDirectory), "Readme")
## Determine if we're supposed to be working in a specific folder, or grabbing the latest
if (!exists("resultsDirectoryName")){
  maxidx <- which.max(as.Date(resultDir))
  rprtDate <- max(as.Date(resultDir), na.rm = TRUE)
  resultsDirectoryName <- paste0(resultDirectory, resultDir[maxidx])
} else{
  rprtDate <- as.Date(resultsDirectoryName)
  resultsDirectoryName <- paste0(resultDirectory, resultsDirectoryName)
}
maxPredDate <- max( rprtDate + 30, as.Date("2020-06-01"))
clean_idx <- ifelse(TEMPLATE_STATE_IDX < 10, 
                    paste0("0", trimws(TEMPLATE_STATE_IDX)), 
                    TEMPLATE_STATE_IDX)
rsltFileIndicator <- grepl(clean_idx, substr(dir(resultsDirectoryName), 1, 2))

if(sum(rsltFileIndicator) > 0){
  resultFiles <- paste0(resultsDirectoryName,"/", 
                        dir(resultsDirectoryName)[rsltFileIndicator])
}else{
  stop("\nResults files were not found\n")
}

analysisResults <- lapply(resultFiles, function(rf){
  e <- new.env()
  load(rf, envir = e)
  e
})
```

```{r,echo=FALSE, results='asis'}
cat(paste0("<h1>Using report compiled on: ", rprtDate, "</h1>\n"))
set.seed(as.numeric(rprtDate))
```

## Introduction

*Some Text on the Models used goes here*


### Reproductive Numbers

Interventions to control infectious diseases strive to drive the "reproductive number" below 1. While this quantity comes in many forms, and is 
derived in many ways for different models, in general it captures the number of secondary infections expected per infectious individual. When 
this number falls below 1, the spread of the disease is expected to slow and stop, while values greater than 1 indicate that the disease is likely to continue to spread exponentially until some other phenomenon interrupts transmission (additional interventions, natural behavioral changes, a lack of susceptible individuals). 

The plot below provides the latest plot estimating our relationship to this threshold. Namely, it estimates the probability that the reproductive number has fallen below 1 at any point during the outbreak in Iowa, and provides a forecast 

### Mortality Projections

The models used here are trained on publicly available mortality data - as such, predicting mortality is one of their principal uses. In this document, we present the latest statewide projections, as well as a comparison to the official May-02 report to see how the predictions have evolved over time (where appropriate).

# {.tabset}

## Latest Results {.tabset}

These results are based on the most recently available models (generally 1-2 days out of date)

### Latest EA-RN information for Iowa 


```{r, echo=FALSE}
R0frame <- Reduce("cbind", lapply(1:length(analysisResults), function(i){
  sims <- analysisResults[[i]]$single_state_results$result$simulationResults
  
  reaOut <- Reduce("cbind", lapply(1:length(sims), function(j){
    REA <- sims[[j]]$R_EA
    REA
  }))
  
}))
## Each eps is the epsilon associated with the data produced from that specific particle (wording)
eps <- Reduce("c", lapply(1:length(analysisResults), function(i){
  analysisResults[[i]]$single_state_results$result$modelObject$epsilon
}))
invEps <- 1/eps
invEps <- invEps/sum(invEps)
sampIdx <- sample(1:length(invEps), size = 1000, prob = invEps, replace = TRUE)
R0samp <- R0frame[,sampIdx]
dtobj <- analysisResults[[1]]$single_state_results$data$date
R0bounds <- apply(R0samp, 1, quantile, probs = c(0.1, 0.25, 0.5,0.75,0.9))
out <- data.frame(date = dtobj, q1 = R0bounds[1,], 
                  q25 = R0bounds[2,],
                  q5 = R0bounds[3,], 
                  q75 = R0bounds[4,],
                  q9 = R0bounds[5,],
                  Observed =  analysisResults[[1]]$single_state_results$data$deaths)
outFilt <- out[ out$date <= maxPredDate,]
p1 <- ggplot(data = outFilt, aes(x=date, y = q5)) + 
  geom_line(color="blue") + 
  geom_vline(xintercept = as.Date("2020-06-01")) + 
  geom_ribbon(aes(ymin = q1, ymax=q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = q25, ymax=q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("R-EA") + xlab("Date") + 
  ggtitle("Reproductive Number")+ 
  ggtitle("Reproductive Number - Current Estimates and Projections\n 50% and 80% Intervals")+  
  geom_hline(yintercept = 1,linetype = "dashed", color = "black", size = 1)+
  theme_light()
p1
```

### Latest Projected Mortality

```{r, echo = FALSE}
## Get mortality projections 
Mortalityframe <- Reduce("cbind", lapply(1:length(analysisResults), function(i){
  sims <- analysisResults[[i]]$single_state_results$result$simulationResults
  
YOut <- Reduce("cbind", lapply(1:length(sims), function(j){
    pi_R <- analysisResults[[i]]$single_state_results$result$params[j,]
    pi_R <- pi_R[names(pi_R) == "report_fraction"]
    deaths <- sims[[j]]$R_star*pi_R
    deaths
  }))
  YOut
}))
eps <- Reduce("c", lapply(1:length(analysisResults), function(i){
  analysisResults[[i]]$single_state_results$result$modelObject$epsilon
}))
invEps <- 1/eps
invEps <- invEps/sum(invEps)
sampIdx <- sample(1:length(invEps), size = 1000, prob = invEps, replace = TRUE)
MortalityframeSamp <- Mortalityframe[,sampIdx]
MortalityframeSamp <- apply(MortalityframeSamp, 2, cumsum)
dtobj <- analysisResults[[1]]$single_state_results$data$date
may1Idx <- which(dtobj == "2020-05-01")
mortalityQuantiles <- apply(MortalityframeSamp, 1, quantile, probs = c(0.1,0.25, 0.5,0.75,0.9))
out <- data.frame(date = dtobj, q1 = mortalityQuantiles[1,], 
                  q25 = mortalityQuantiles[2,],
                  q5 = mortalityQuantiles[3,], 
                  q75 = mortalityQuantiles[4,],
                  q9 = mortalityQuantiles[5,],
                  Observed =  analysisResults[[1]]$single_state_results$data$deaths)
outFilt <- out[out$date >= as.Date("2020-04-01") &
                 out$date <= maxPredDate,]
outFilt.latest <- outFilt
p1 <- ggplot(data = outFilt, aes(x=date, y = q5)) + 
  geom_vline(xintercept = as.Date("2020-06-01")) +
  geom_point(x = outFilt$date, y = outFilt$Observed) + 
  geom_line(color="blue") + 
  geom_ribbon(aes(ymin = q1, ymax=q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = q25, ymax=q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("Cumulative Mortality") + xlab("Date") + 
  ggtitle("Six Week Cumulative Mortality")+ 
 
  ggtitle("Projected Cumulative Mortality\n 50% and 80% Intervals") +  
  theme_light()
suppressWarnings(
  print(p1)
)
p2 <- ggplot(data = outFilt, aes(x=date, y = q5))+  scale_y_continuous(trans='log10') + 
  geom_line(color="blue") + 
  geom_vline(xintercept = as.Date("2020-06-01")) +
  geom_point(x = outFilt$date, y = log10(outFilt$Observed) )+
  geom_ribbon(aes(ymin = outFilt$q1, ymax=outFilt$q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = outFilt$q25, ymax=outFilt$q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("Cumulative Mortality") + xlab("Date") + 
  ggtitle("Projected Cumulative Mortality\n (log10 scale)")+  
  theme_light() 
suppressWarnings(
  print(p2)
)
out.joined <- out  %>% mutate(CIprnt = paste0(round(q5), " (", round(q1), ", ", round(q9), ")"))
out.ToPrint <- select(out.joined, date, Observed, CIprnt) %>% filter(date <= maxPredDate)
colnames(out.ToPrint) <- c("Date", "Observed Mortality", "Predicted Mortality")
kable(out.ToPrint, digits = 1,align = c("c"), caption = "Mortality: Latest") %>% 
  kable_styling(bootstrap_option = c("striped", "condensed"), 
                full_width = FALSE) 
 
```
### Latest Cumulative Case Count

```{r, echo = FALSE}
## Get Incidence projections 
Incidenceframe <- Reduce("cbind", lapply(1:length(analysisResults), function(i){
  sims <- analysisResults[[i]]$single_state_results$result$simulationResults
  YOut <- Reduce("cbind", lapply(1:length(sims), function(j){
    ## I don't think I need report fraction for latent variable?
      pi_R <- analysisResults[[i]]$single_state_results$result$params[j,]
      pi_R <- pi_R[names(pi_R) == "report_fraction"]
      cases <- sims[[j]]$I_star
      cases
    }))
  YOut
}))


eps <- Reduce("c", lapply(1:length(analysisResults), function(i){
  analysisResults[[i]]$single_state_results$result$modelObject$epsilon
}))
invEps <- 1/eps
invEps <- invEps/sum(invEps)
sampIdx <- sample(1:length(invEps), size = 1000, prob = invEps, replace = TRUE)
IncidenceframeSamp <- Incidenceframe[,sampIdx]
IncidenceframeSamp <- apply(IncidenceframeSamp, 2, cumsum)
dtobj <- analysisResults[[1]]$single_state_results$data$date
may1Idx <- which(dtobj == "2020-05-01")
IncidenceQuantiles <- apply(IncidenceframeSamp, 1, quantile, probs = c(0.1,0.25, 0.5,0.75,0.9))
out <- data.frame(date = dtobj, 
                  q1 = IncidenceQuantiles[1,], 
                  q25 = IncidenceQuantiles[2,],
                  q5 = IncidenceQuantiles[3,], 
                  q75 = IncidenceQuantiles[4,],
                  q9 = IncidenceQuantiles[5,],
                  Observed =  analysisResults[[1]]$single_state_results$data$cases)
outFilt <- out[out$date >= as.Date("2020-04-01") &
                 out$date <= maxPredDate,]
outFilt.latest <- outFilt
p1 <- ggplot(data = outFilt, aes(x=date, y = q5)) + 
  geom_vline(xintercept = as.Date("2020-06-01")) +
  geom_point(x = outFilt$date, y = outFilt$Observed) + 
  geom_line(color="blue") + 
  geom_ribbon(aes(ymin = q1, ymax=q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = q25, ymax=q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("Cumulative Incidence") + xlab("Date") + 
  ggtitle("Six Week Cumulative Cases")+ 
 
  ggtitle("Projected Cumulative Cases\n 50% and 80% Intervals") +  
  theme_light()
suppressWarnings(
  print(p1)
)
p2 <- ggplot(data = outFilt, aes(x=date, y = q5))+  scale_y_continuous(trans='log10') + 
  geom_line(color="blue") + 
  geom_vline(xintercept = as.Date("2020-06-01")) +
  geom_point(x = outFilt$date, y = log10(outFilt$Observed) )+
  geom_ribbon(aes(ymin = outFilt$q1, ymax=outFilt$q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = outFilt$q25, ymax=outFilt$q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("Cumulative Incidence") + xlab("Date") + 
  ggtitle("Projected Cumulative Incidence\n (log10 scale)")+  
  theme_light() 
suppressWarnings(
  print(p2)
)
out.joined <- out  %>% mutate(CIprnt = paste0(round(q5), " (", round(q1), ", ", round(q9), ")"))
out.ToPrint <- select(out.joined, date, Observed, CIprnt) %>% filter(date <= maxPredDate)
colnames(out.ToPrint) <- c("Date", "Observed Incidence", "Predicted Incidence")
kable(out.ToPrint, digits = 1,align = c("c"), caption = "Incidence: Latest") %>% 
  kable_styling(bootstrap_option = c("striped", "condensed"), 
                full_width = FALSE) 
 
```



## Comparison to May 02 Report {.tabset}
 
These results are reproduced from the models included in the May 02 report using model M2. The mortality projections are compared to the
subsequently reported counts. 

### Reproductive Number Projections 

```{r, echo=FALSE,results='asis'}
if (rprtDate == as.Date("2020-05-02")){
  cat("<strong>Comparison omitted - report date is the same as the comparison report.</strong>") 
}else if (rprtDate > as.Date("2020-06-01")){
  cat("<strong>Comparison omitted - report date is past June 01 and comparison is not meaningful.</strong>") 
} else if (rprtDate <= as.Date("2020-06-01")){
  
rwd <- paste0(resultDirectory, "2020-05-02/results/")
resultFiles <- paste0(rwd,"/", dir(rwd)[grepl("iowa", dir(rwd))])
analysisResults <- lapply(resultFiles, function(rf){
  e <- new.env()
  load(rf, envir = e)
  e
})
R0frame <- Reduce("cbind", lapply(1:length(analysisResults), function(i){
  sims <- analysisResults[[i]]$single_state_results$result$simulationResults
  
  reaOut <- Reduce("cbind", lapply(1:length(sims), function(j){
    REA <- sims[[j]]$R_EA
    REA
  }))
  
}))
eps <- Reduce("c", lapply(1:length(analysisResults), function(i){
  analysisResults[[i]]$single_state_results$result$modelObject$epsilon
}))
invEps <- 1/eps
invEps <- invEps/sum(invEps)
sampIdx <- sample(1:length(invEps), size = 1000, prob = invEps, replace = TRUE)
R0samp <- R0frame[,sampIdx]
dtobj <- analysisResults[[1]]$single_state_results$data$date
dtIdx <- max(which(is.finite(analysisResults[[1]]$single_state_results$result$modelObject$modelComponents$data_model$Y)))
R0bounds <- apply(R0samp, 1, quantile, probs = c(0.1, 0.25, 0.5,0.75,0.9))
out <- data.frame(date = dtobj, q1 = R0bounds[1,], 
                  q25 = R0bounds[2,],
                  q5 = R0bounds[3,], 
                  q75 = R0bounds[4,],
                  q9 = R0bounds[5,],
                  Observed =  analysisResults[[1]]$single_state_results$data$deaths)
outFilt <- out[out$date >= as.Date("2020-04-01") &
                 out$date <= as.Date("2020-06-01"),]
p1 <- ggplot(data = outFilt, aes(x=date, y = q5)) + 
  geom_line(color="blue") + 
  geom_vline(xintercept = as.Date("2020-06-01")) +
  geom_ribbon(aes(ymin = q1, ymax=q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = q25, ymax=q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("R-EA") + xlab("Date") + 
  ggtitle("Reproductive Number")+ 
  ggtitle("Reproductive Number - Status Quo\n 50% and 80% Intervals")+  
  geom_hline(yintercept = 1,linetype = "dashed", color = "black", size = 1)+
  theme_light()
suppressWarnings({
  print(p1)
})
}
```

### Mortality Projections


```{r,echo=FALSE,results='asis'}
## Get mortality projections 
if (rprtDate == as.Date("2020-05-02")){
  cat("<strong>Comparison omitted - report date is the same as the comparison report.</strong>") 
}else if (rprtDate > as.Date("2020-06-01")){
  cat("<strong>Comparison omitted - report date is past June 01 and comparison is not meaningful.</strong>") 
} else if (rprtDate <= as.Date("2020-06-01")){
  
Mortalityframe <- Reduce("cbind", lapply(1:length(analysisResults), function(i){
  sims <- analysisResults[[i]]$single_state_results$result$simulationResults
  
YOut <- Reduce("cbind", lapply(1:length(sims), function(j){
    pi_R <- analysisResults[[i]]$single_state_results$result$params[j,]
    pi_R <- pi_R[names(pi_R) == "report_fraction"]
    deaths <- sims[[j]]$R_star*pi_R
    deaths
  }))
  YOut
}))
eps <- Reduce("c", lapply(1:length(analysisResults), function(i){
  analysisResults[[i]]$single_state_results$result$modelObject$epsilon
}))
invEps <- 1/eps
invEps <- invEps/sum(invEps)
sampIdx <- sample(1:length(invEps), size = 1000, prob = invEps, replace = TRUE)
MortalityframeSamp <- Mortalityframe[,sampIdx]
MortalityframeSamp <- apply(MortalityframeSamp, 2, cumsum)
dtobj <- analysisResults[[1]]$single_state_results$data$date
may1Idx <- which(dtobj == "2020-05-01")
mortalityQuantiles <- apply(MortalityframeSamp, 1, quantile, probs = c(0.1,0.25, 0.5,0.75,0.9))
out <- data.frame(date = dtobj, q1 = mortalityQuantiles[1,], 
                  q25 = mortalityQuantiles[2,],
                  q5 = mortalityQuantiles[3,], 
                  q75 = mortalityQuantiles[4,],
                  q9 = mortalityQuantiles[5,],
                  Observed =  analysisResults[[1]]$single_state_results$data$deaths)
outFilt <- out[out$date >= as.Date("2020-04-01") & 
                out$date <= as.Date("2020-06-01"),]
outFilt.latest.sub <- filter(outFilt.latest, date %in% outFilt$date)
if (!all(outFilt.latest.sub$date == outFilt$date)){
  stop("Date Comparison is broken")
}
pred <- is.na(outFilt$Observed) & !is.na(outFilt.latest.sub$Observed)
outFilt$Observed <- outFilt.latest.sub$Observed
cols <- c("black", "red")
p1 <- ggplot(data = outFilt, aes(x=date, y = q5)) + 
  geom_vline(xintercept = as.Date("2020-06-01")) +
  geom_line(color="blue") + 
  geom_vline(xintercept = outFilt$date[min(which(pred))-1]) +
  geom_ribbon(aes(ymin = outFilt$q1, ymax=outFilt$q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = outFilt$q25, ymax=outFilt$q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("Cumulative Mortality") + xlab("Date") + 
  ggtitle("Six Week Cumulative Mortality")+ 
  geom_point(x = outFilt$date, y = outFilt$Observed, color = cols[pred+1]) + 
  ggtitle("Projected Cumulative Mortality\n 50% and 80% Intervals")+
  theme_light()
p2 <- ggplot(data = outFilt, aes(x=date, y = q5))+  scale_y_continuous(trans='log10') + 
  geom_line(color="blue") +
  geom_vline(xintercept = outFilt$date[min(which(pred))-1]) +
  geom_point(x = outFilt$date, y = log10(outFilt$Observed) , color = cols[pred+1])+
  geom_ribbon(aes(ymin = outFilt$q1, ymax=outFilt$q9), 
              linetype = 2, alpha = 0.05, size = 2) +
  geom_ribbon(aes(ymin = outFilt$q25, ymax=outFilt$q75), 
              linetype = 2, alpha = 0.15, size = 2) +
  ylab("Cumulative Mortality") + xlab("Date") + 
  ggtitle("Projected Cumulative Mortality\n (log10 scale)")+  
  theme_light() 
suppressWarnings({
  print(p1)
  print(p2)
})
out.joined <- left_join(out, mutate(outFilt, latestY = Observed) %>% select(date, latestY), by = "date") %>%
  mutate(ObservedPrnt = ifelse(is.na(Observed) & !is.na(latestY), paste0(latestY, "*"), Observed)) %>%
  mutate(CIprnt = paste0(round(q5), " (", round(q1), ", ", round(q9), ")"))
out.ToPrint <- select(out.joined, date, ObservedPrnt, CIprnt) %>% filter(date <= as.Date("2020-06-01"))
colnames(out.ToPrint) <- c("Date", "Observed Mortality", "Predicted Mortality")
  
}
```




```{r, echo = FALSE}
# Show mortality table if valid
if (rprtDate > as.Date("2020-05-02") && rprtDate <= as.Date("2020-06-01")){
  kable(out.ToPrint, digits = 1,align = c("c"), caption = "Mortality: Latest") %>% 
  kable_styling(bootstrap_option = c("striped", "condensed"), 
                full_width = FALSE) 
}
```